---
title: "Untitled"
author: "Arefeva Anna"
date: "23 12 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(readxl)
library(tidymodels)
library(embed)
library(themis)
library(ggcorrplot)
library(psych)
library(yardstick)
library(vip)
library(Boruta)
library(rpart)
library(uwot)
library(plotly)
library(cluster)
library(dplyr)
## input_data
#reading individual pages of a file with bacterial function and leading them to identical form - a column indicating functions and 1/0 values for each bacterial name

bac_gr_funcs1 <- read_excel("Bacterial group functions.xlsx", sheet = 'Патогены и нежелательные')

bac_gr_funcs1 <- bac_gr_funcs1 %>%
  mutate(Condicionally_norm = ifelse(Bacteria_category == 'Условно-нормальный', 1, 0),
         Patogen = ifelse(Bacteria_category == 'Патоген', 1, 0),
          TaxonName = gsub(" ", "_", TaxonName),
         Bacteria_tax = paste(TaxonName, Rank, sep = "_"))%>%
  select(-TaxonName, -Rank, -Bacteria_category)



bac_gr_funcs2 <- read_excel("Bacterial group functions.xlsx", sheet = 'Продуценты серотонина')

bac_gr_funcs2 <- bac_gr_funcs2 %>%
  distinct(.keep_all = TRUE)%>%
  mutate(Value = 1) %>%
  pivot_wider(names_from = Neuromediator, values_from = Value)%>%
  mutate (TaxonName = gsub(" ", "_", TaxonName),
         Bacteria_tax = paste(TaxonName, Rank, sep = "_"))%>%
  select(-TaxonName, -Rank)



bac_gr_funcs3 <- read_excel("Bacterial group functions.xlsx", sheet = 'Пробиотики') 
bac_gr_funcs3 <- bac_gr_funcs3 %>%
  mutate(Probiotics = 1,
          TaxonName = gsub(" ", "_", TaxonName),
         Bacteria_tax = paste(TaxonName, Rank, sep = "_"))%>%
  select(-TaxonName, -Rank)



bac_gr_funcs4 <- read_excel("Bacterial group functions.xlsx", sheet = 'Бактерии с особыми свойстами') 



bac_gr_funcs4 <- bac_gr_funcs4 %>%
  mutate(Specific = 1,
          TaxonName = gsub(" ", "_", TaxonName),
         Bacteria_tax = paste(TaxonName, Rank, sep = "_"))%>%
  select(-TaxonName, -Rank)


bac_gr_funcs5 <- read_excel("Bacterial group functions.xlsx", sheet = 'Витаминные') 

bac_gr_funcs5 <- bac_gr_funcs5 %>%
  mutate(Value = 1) %>%
  pivot_wider(names_from = Vitamin, values_from = Value)%>%
  mutate(TaxonName = gsub(" ", "_", TaxonName),
         Bacteria_tax = paste(TaxonName, Rank, sep = "_"))%>%
  select(-TaxonName, -Rank)


bac_gr_funcs6 <- read_excel("Bacterial group functions.xlsx", sheet = 'Продуценты КЦЖК') 
bac_gr_funcs6 <- bac_gr_funcs6 %>%
  mutate(TaxonName = gsub(" ", "_", TaxonName),
         Bacteria_tax = paste(TaxonName, Rank, sep = "_"))%>%
  select(-TaxonName, -Rank)



bac_gr_funcs7<- read_excel("Bacterial group functions.xlsx", sheet = 'Вредные привычки') 

bac_gr_funcs7 <- bac_gr_funcs7 %>%
  distinct(.keep_all = TRUE)%>%
  pivot_wider(names_from = Habbit, values_from = Habit_state)%>%
  mutate(TaxonName = gsub(" ", "_", TaxonName),
         Bacteria_tax = paste(TaxonName, Rank, sep = "_"))%>%
  select(-TaxonName, -Rank)
```


```{r setup, include=FALSE}
#merging pages with bacterial functions into one table

bac_gr_funcs <- full_join(bac_gr_funcs1, bac_gr_funcs2)
bac_gr_funcs <- full_join(bac_gr_funcs, bac_gr_funcs3)

bac_gr_funcs <- full_join(bac_gr_funcs, bac_gr_funcs4)
bac_gr_funcs <- full_join(bac_gr_funcs, bac_gr_funcs5)
bac_gr_funcs <- full_join(bac_gr_funcs, bac_gr_funcs6)
bac_gr_funcs <- full_join(bac_gr_funcs, bac_gr_funcs7)


#file with data of patients 

fbh <- read.csv('final_bacteria_health.csv') 
fb140 <- read.csv("final_bacteria_ibs_140.csv")                         
fbh_stats <- openxlsx::read.xlsx("final_health_statistic.xlsx")          
fb140_stats <- openxlsx::read.xlsx("final_ibs_140_statistic.xlsx")  

fb140_long <- fb140 %>%
  pivot_longer(cols = X01D2Z36_F:possible.genus.03_G,
               names_to = 'Bacteria_tax',
               values_to = 'Value')


fbh_long <- fbh %>%
  pivot_longer(cols = X01D2Z36_F:possible.genus.03_G,
               names_to = 'Bacteria_tax',
               values_to = 'Value')

fb140_mrg <- fb140_long %>% 
  inner_join(fb140_stats, by="patient_ID")%>%
  select(-Main_Disease)

fbh_mrg <- fbh_long %>% 
  inner_join(fbh_stats, by="patient_ID")

fb_long <- bind_rows(fbh_mrg, fb140_mrg)

#full dataset

fb_all <- fb_long %>% 
  full_join(bac_gr_funcs, by="Bacteria_tax")
#возраст
fb_all <- fb_all %>%
  mutate(Age = case_when (Age == ' ' ~ (Seq_date-Birth_Year), 
                          Age != ' ' ~ Age))

#adding Age_group (WHO classification)

fb_all <- fb_all%>%
  mutate(Young = case_when(Age <= 44 | Age_max <= 44 | Age_min <= 44 ~ 1, 
                           Age > 44 | Age_min > 44  ~ 0),
         Adult = case_when ((Age >= 45 & Age <= 60) | (Age_max >= 45 &Age_max <= 60) | (Age_min >= 45 & Age_min <= 60) ~ 1, 
                            Age < 44 | Age > 60 | Age_max > 60 | Age_max < 44 | Age_min < 44 | Age_min > 60 ~ 0),
         Old = case_when (Age > 60 | Age_max > 60 ~1,
                          Age <= 60 | Age_max <= 60 ~ 0)
  )
         

#BMI

fb_all <- fb_all %>%
  mutate(BMI_min = case_when (BMI_min == ' '~(Weight_min/Height_min/Height_min*10000), BMI_min != ' ' ~ BMI_min),
         BMI_max = case_when(BMI_max == ' ' ~ (Weight_max/Height_max/Height_max*10000), BMI_max  != ' ' ~ BMI_max),
         BMI = Weight_kg/Height_cm/Height_cm*10000)

summary(fb_all)

fb_all <- fb_all %>%
  mutate(Thin = case_when (BMI < 18.5 | BMI_min < 18.5 |BMI_max < 18.5 ~ 1, BMI >18.5 | BMI_min > 18.5 ~ 0 ),
         Norm_weight = case_when (BMI >=18.5 & BMI < 25.0 | BMI_min>=18.5&BMI_min<25.0 | BMI_max>= 18.5~1, BMI <18.5 | BMI >= 25.0 |BMI_max < 18.5 | BMI_min>=25.0 ~0 ),
         Fat = case_when(BMI >=25.0 |BMI_max >=25.0 ~ 1, BMI < 25.0 | BMI_max < 25.0 ~ 0))


#counting missing values

NA_count <- fb_all %>%
    mutate(n = n()) %>% 
    summarise(across(everything(), ~ sum(is.na(.x))),
              n = n()) %>% 
    select(n, where(function(x) x != 0)) %>% 
    mutate(across(everything()), round(. / n, 4)) %>% 
    select(-n) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "Feature",
      values_to = "Value"
    ) %>% 
    arrange(desc(Value))



#saving dataset
data.table::fwrite(fb_all, "all_long.csv")

data.table::fwrite(all_data_O, 'all_data_O.csv')


```

### Order
```{r setup, include=FALSE}
#select data by Order (_O)

all_data_O <- fb_all %>%
  mutate(Bacteria_tax = ifelse(grepl("_O", fb_all$Bacteria_tax) == TRUE, Bacteria_tax, 'NO'))
all_data_O <- all_data_O[all_data_O$Bacteria_tax != 'NO', ]


#NA in Order

NA_count_O <- all_data_O %>%
    mutate(n = n()) %>% 
    summarise(across(everything(), ~ sum(is.na(.x))),
              n = n()) %>% 
    select(n, where(function(x) x != 0)) %>% 
    mutate(across(everything()), round(. / n, 4)) %>% 
    select(-n) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "Feature",
      values_to = "Value"
    ) %>% 
    arrange(desc(Value))


#exclusion of bacteria with bacterial load = 0 in >280 patients

bac_O <- all_data_O %>%
  select(Bacteria_tax, Value)%>%
  count(Bacteria_tax, Value == 0)

bac_O <- bac_O%>%
  rename(Value = 'Value == 0')

bac_O <- bac_O[bac_O$Value == TRUE & bac_O$n <= 280, ]

bac_O_names <- bac_O$Bacteria_tax
bac_O_names <- na.omit(bac_O_names)
bac_O_names

all_data_O_new <- all_data_O %>%
  filter(.$Bacteria_tax %in% bac_O_names )

#removing variables with 99-100% NA

data <- all_data_O_new%>%
  select(-c(Instrument:Seq_date), -c(Birth_Year:BMI_max), -c(Weight_kg:Алкоголь), -Drugs, -BMI)

NA_count_data <- data %>%
    mutate(n = n()) %>% 
    summarise(across(everything(), ~ sum(is.na(.x))),
              n = n()) %>% 
    select(n, where(function(x) x != 0)) %>% 
    mutate(across(everything()), round(. / n, 4)) %>% 
    select(-n) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "Feature",
      values_to = "Value"
    ) %>% 
    arrange(desc(Value))

#descriptive statistics 
summary(data)

#descriptive statistics for quantitative variables (grouping by bacteria category)
statistics <- list(
  `_n` = ~length(.x) %>% as.character(),
  `_NA` = ~sum(is.na(.x)) %>% as.character(),
  `_Mean` = ~ifelse(sum(!is.na(.x)) == 0, "Н/П*", mean(.x, na.rm = TRUE) %>% round(2) %>% as.character()),
  `_SD` = ~ifelse(sum(!is.na(.x)) < 3, "Н/П*", sd(.x, na.rm = TRUE) %>% round(2) %>% as.character()),
  `_Min` = ~ifelse(sum(!is.na(.x)) == 0, "Н/П*", min(.x, na.rm = TRUE) %>% round(2) %>% as.character()),
  `_Max` = ~ifelse(sum(!is.na(.x)) == 0, "Н/П*", max(.x, na.rm = TRUE) %>% round(2) %>% as.character()),
  `_Median` = ~ifelse(sum(!is.na(.x)) == 0, "Н/П*", median(.x, na.rm = TRUE) %>% round(2) %>% as.character()
))

group_data_O_stat <- all_data_O %>%
  select(Bacteria_tax, where(is.numeric)) %>%
  group_by(Bacteria_tax) %>%
  summarize(across(where(is.numeric), statistics)) %>% 
  pivot_longer(!Bacteria_tax) %>%
  separate(name, into = c('Переменная', 'Статистика'), sep = "__") %>%
  rename('Значение' = value) %>%
  print(., n=Inf)
```


```{r setup, include=FALSE}
### batch-effect ~ research_ID

kruskal_res <-data %>% split(~ Bacteria_tax) %>% map(function(subdata) kruskal.test(Value ~ research_ID, data = subdata)$p.value)%>% enframe() %>% unnest(cols = value) 

kruskal_res <- kruskal_res %>%
  ungroup()%>%
  select(Bacteria_tax, p.value)%>%
  arrange(p.value)


kruskal_res <- kruskal_res %>%
  mutate(p.adjusted = p.adjust(p.value, method = 'BH'))

significant <- kruskal_res %>%
  filter(p.adjusted < 0.05)


#visualisation
data %>% 
  filter(Value != 0 & research_ID != 0) %>% 
  ggplot(aes(x=research_ID, y=Value)) + 
    geom_point(size=1)
```


```{r setup, include=FALSE}
#search for significant differences between sick and healthy people
test_res <-data %>% split(~ Bacteria_tax) %>% map(function(subdata) wilcox.test(Value ~ Health_state, data = subdata)$p.value)%>% enframe() %>% unnest(cols = value) 


  showbac.sex <- data %>%
    filter(Bacteria_tax %in% test_res$name) %>% 
    select(Bacteria_tax, Value, Health_state) %>% 
    group_by(Health_state, Bacteria_tax) %>% 
    summarise(meanValue = mean(Value)*100) %>% 
    filter(meanValue > 0) %>% 
    arrange(meanValue) %>% 
    filter(meanValue > quantile(meanValue, 0.05) & meanValue < quantile(meanValue, 0.95)) %>% 
    ungroup() %>% 
    rename(word = Bacteria_tax, freq = meanValue) %>% 
    mutate(freq = ceiling(freq)) 
  
  showbac.health <- showbac.sex %>%
    filter(Health_state == "Health") %>%
    select(-Health_state) %>% 
    arrange(desc(freq)) %>% 
    as.data.frame()
  
  showbac.desease <- showbac.sex %>%
    filter(Health_state == "Desease") %>%
    select(-Health_state) %>% 
    arrange(desc(freq)) %>% 
    as.data.frame()


#search for significant differences between male and female
data3 <- data [data$Sex == 'male'|data$Sex == 'female', ]
test_res1 <-data3 %>% split(~ Bacteria_tax) %>% map(function(subdata) wilcox.test(Value ~ Sex, data = subdata)$p.value)%>% enframe() %>% unnest(cols = value)

test_res1 <- test_res1 %>%
  mutate(p.adjusted = p.adjust(value, method = 'BH'))

significant_3 <- test_res1 %>%
  filter(p.adjusted < 0.05)

#visualisation (word cloud) for bacterial taxon between male and female
  showbac.sex <- data3 %>%
    filter(Bacteria_tax %in% test_res1$name) %>% 
    select(Bacteria_tax, Value, Sex) %>% 
    group_by(Sex, Bacteria_tax) %>% 
    summarise(meanValue = mean(Value)*100) %>% 
    filter(meanValue > 0) %>% 
    arrange(meanValue) %>% 
    filter(meanValue > quantile(meanValue, 0.05) & meanValue < quantile(meanValue, 0.95)) %>% 
    ungroup() %>% 
    rename(word = Bacteria_tax, freq = meanValue) %>% 
    mutate(freq = ceiling(freq)) 
  
  showbac.sex_m <- showbac.sex %>%
    filter(Sex == "male") %>%
    select(-Sex) %>% 
    arrange(desc(freq)) %>% 
    as.data.frame()
  
  showbac.sex_f <- showbac.sex %>%
    filter(Sex == "female") %>%
    select(-Sex) %>% 
    arrange(desc(freq)) %>% 
    as.data.frame()


```

### ML
```{r setup, include = FALSE, fig.width=9, fig.height=8, fig.showtext=TRUE, echo=FALSE}
data %>% glimpse()

data1 <-data%>%
  pivot_wider(names_from = Bacteria_tax, values_from = Value)%>%
  mutate(across(c(X0319.6G20_O:eub62A3_O), ~ as.numeric(.x)))%>%
   mutate(across(where(is.character), as.factor))

data1 %>% glimpse()


data1 %>% count(Health_state)


```

### pre-processing

```{r setup, include = FALSE}

data1<-data1 %>%
  mutate(Race = case_when(Race != NA & Race !='Caucasian' ~ 'Other'))%>%
   mutate(across(where(is.character), as.factor))


data1 %>%
  select(where(is.factor)) %>%
  map(function(x) sum(is.na(x)) / length(x)) %>%
  enframe() %>%
  unnest(cols = value) %>%
  arrange(desc(value))

data1 %>%
  select(where(is.factor)) %>%
  map(table)




```



```{r setup, include = FALSE, , fig.width=9, fig.height=35, fig.showtext=TRUE, echo=FALSE}
#removing uninformative variables
data2 <- data1 %>%
  select (-patient_ID, -research_ID, -c(Sex:Sleep_duration), -c(Thin:Fat))%>%
  drop_na()%>%
  mutate(across(where(is.character), as.factor))

data2 %>%
  select(where(is.factor)) %>%
  map(function(x) sum(is.na(x)) / length(x)) %>%
  enframe() %>%
  unnest(cols = value) %>%
  arrange(desc(value))


#Boruta for highlighting significant variables
data2%>%
  glimpse()
Boruta(Health_state ~ ., data2, ntree = 500, maxRuns = 300) %>%
  TentativeRoughFix() -> boruta_trained
boruta.sig.vars <- boruta_trained %>%
  attStats() %>%
  filter(decision == "Confirmed" & meanImp > 5) %>% 
  arrange(desc(meanImp)) %>% 
  rownames_to_column(var = "ImpVar") %>% 
  select(ImpVar) %>% 
  pull()

boruta_trained %>%
  attStats() %>%
  rownames_to_column("Переменная") %>%
  mutate(`Переменная` = `Переменная` %>% fct_reorder(`meanImp`)) %>%
  
  ggplot(aes(y = `Переменная`, x = meanImp, colour = decision)) +
  geom_point() +
  geom_errorbar(aes(xmin = minImp, xmax = maxImp, width = 0.1)) +
  xlab("Среднее снижение энтропии") +
  labs(color = "Значимость переменной") +
  theme(legend.position = "bottom")
  #scale_color_disrete(labels = c("Подтверждённая", "Неподтверждённая"))


```


```{r setup, include = FALSE}


CPK <- data2[sample(nrow(data2)), ]
CPK <- CPK %>% select(Health_state, boruta.sig.vars)%>%
  mutate(Health_state = ifelse(Health_state == "Disease", 1, 0))

  
split_train_test <- initial_split(data2, strata = Health_state, prop = 0.9)

CPK_train <- split_train_test %>% training()
CPK_test <- split_train_test %>% testing()



```



```{r setup, include = FALSE}
cat_metric <- yardstick::metric_set(
  
    yardstick::bal_accuracy,
    yardstick::precision,
    yardstick::recall,
    yardstick::f_meas,
    yardstick::sensitivity,
    yardstick::specificity,
    yardstick::j_index
  
  )
```
### creating a recipe

```{r setup, include = FALSE}
CPK_recipe <- recipe(Health_state ~ ., CPK_train) %>%
  
  step_impute_bag(all_numeric_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_nzv(all_predictors()) %>%
  step_lincomb(all_numeric_predictors()) %>%
  #step_normalize(all_numeric_predictors()) %>%
  step_lencode_glm(all_nominal_predictors(), outcome = vars(Health_state)) %>%
  step_adasyn(Health_state) #%>%
  #prep()

#heart_recipe %>% juice() %>% count(HeartDisease)
#heart_train %>% count(HeartDisease)

```

### definition of model

```{r setup, include = FALSE}
rf_model <- rand_forest(mode = "classification", mtry = tune(), trees = tune(), min_n = tune()) %>%
  set_engine("ranger")
```

### definition of samples
```{r setup, include = FALSE}
cv_samples <- vfold_cv(CPK_train, strata = Health_state, v = 4)
```


### definition of parameters_grid
```{r setup, include = FALSE}
parameters_grid <- grid_max_entropy(mtry(range = c(3, 5)), trees(), min_n(), size = 10)
```


### pipeline
```{r setup, include = FALSE}
reg_workflow <- workflow() %>%
  add_recipe(CPK_recipe) %>%
  add_model(rf_model)

round(19.48*2)/2
```


### learning
```{r setup, include = FALSE}
grid_search <- reg_workflow %>%
  
  tune_grid(
    
    object = reg_workflow,
    resamples = cv_samples,
    grid = parameters_grid,
    control = control_grid(save_pred = TRUE),
  

    metrics = metric_set(sensitivity, specificity, j_index)

)
```

### visualisation
```{r setup, include = FALSE}
grid_search %>%
  collect_metrics() %>%
  ggplot(aes(trees, mean, color = .metric)) +
  geom_errorbar(aes(ymin = mean - std_err, ymax = mean + std_err), alpha = 0.8) +
  geom_line(size = 1)

```


### selection of the best model
```{r setup,  include = FALSE }
best_by_j_index <- grid_search %>% select_best("j_index")

best_by_j_index %>%   flextable::flextable() %>% flextable::theme_box()

final_reg_model <- finalize_workflow(
  
  reg_workflow,
  best_by_j_index
  
)
```

### comparison of metrics on test and training data
```{r setup,  include = FALSE }

final_reg_model %>%
  fit(CPK_train) %>%
  extract_fit_parsnip() %>%
  vi() %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
  geom_col() +
  labs(y = NULL)


final_reg_model %>%
  keras::fit(CPK_train) %>%
  predict(CPK_test) %>%
  pull() -> final_test_prediction

metrics_on_test <- cat_metric(truth = truth_values, estimate = estimate_values, tibble(truth_values = CPK_test$Health_state, estimate_values = final_test_prediction)) %>% rename(test_estimate = `.estimate`) %>% select(!`.estimator`)

# train data

final_reg_model %>%
  keras::fit(CPK_train) %>%
  predict(CPK_train) %>%
  pull() -> final_train_prediction

metrics_on_train <- cat_metric(truth = truth_values, estimate = estimate_values, tibble(truth_values = CPK_train$Health_state, estimate_values = final_train_prediction)) %>% rename(train_estimate = `.estimate`) %>% select(!`.estimator`)

# binding

metrics_on_test %>%
  left_join(metrics_on_train, by = ".metric") %>%
  mutate(differencies = train_estimate - test_estimate) %>% 
  flextable::flextable() %>% flextable::theme_box()
```

### final estimation

```{r setup, include = FALSE}
last_fit(
  
  final_reg_model,
  split_train_test
  
) -> final_model_all_data

final_model_all_data %>%
  extract_workflow() %>%
  predict(CPK, type = "class") %>%
  pull() -> class_prediction

final_reg_model %>%
  extract_workflow() %>%
  predict(CPK, type = "prob") %>%
  pull() -> prob_prediction

log_model_results <- tibble(truth = CPK$Health_state,
                            estimate = class_prediction,
                            prob_yes = 1 - prob_prediction)

cat_metric(truth = truth, estimate = estimate, log_model_results) 

log_model_results %>%
  roc_curve(truth = truth, prob_yes) %>%
  autoplot()
```

# DBSCAN

```{r setup, include = FALSE}


data -> cluster_data_for_vizualisation

data %>% select(!Health_state) -> cluster_data_for_model

recipe(~ ., data = cluster_data_for_model) %>%
  step_scale(all_numeric_predictors()) %>%
  #step_dummy(Sex) %>%
  prep(cluster_data_for_model) %>%
  juice() -> data_recoded

data_recoded %>%
  daisy(metric = "gower") %>%
  umap(n_components = 7) -> umap_data

data_recoded %>%
  daisy(metric = "gower") %>%
  umap(n_components = 3) -> umap_data_for_visualization

umap_data %>%
  daisy(metric = "euclidean") -> dist_data

#dist_data %>%
#  dbscan::dbscan(eps = mean(dist_data)) -> cluster_data_for_model

dist_data %>%
  dbscan::hdbscan(minPts = 5) -> cluster_data

umap_data_for_visualization %>%
  as_tibble() %>%
  mutate(`Кластер` = as.factor(str_c("Кластер №", cluster_data$cluster)),
         Category = cluster_data_for_vizualisation$Category) %>%
  rename(`Компонент 1` = V1,
         `Компонент 2` = V2,
         `Компонент 3` = V3) -> data_3d


plot_ly(data_3d,
        x =~ `Компонент 1`,
        y =~ `Компонент 2`,
        z =~ `Компонент 3`,
        color =~ `Кластер`, hoverinfo = "text", text = ~paste(Category)) %>%
  
  add_markers(size = 1.5)
