---
title: "CPK2_Family"
author: "Stanislav Legkovoy"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(glmnet)
library(Matrix)
library(wordcloud)
library(ggcorrplot)
library(embed)
library(themis)
library(psych)
library(yardstick)
library(vip)
library(Boruta)
library(glmmTMB)
library(rpart)
library(uwot)
library(plotly)
library(cluster)
library(flexclust)
library(factoextra)
library(rattle)
library(brunnermunzel)
```


## Input data
```{r data}

#Bacterial group functions

dt_vec <- c(NULL)

for (i in 1:7) {
  suf <- paste0("bac_gr_funcs", i)
  dt_vec <<- append(suf, dt_vec)
  assign(suf, openxlsx::read.xlsx("./data/Bacterial group functions.xlsx", sheet = i))
}
  
  for (dt in dt_vec) {
    tmp <- get(dt)
    tmp <- filter(tmp, Rank=="F")
    assign(dt, tmp)
  }

rm(tmp)
  
  fbh <- read.csv("./data/final_bacteria_health.csv")                           #final_bacteria_health
  fb140 <- read.csv("./data/final_bacteria_ibs_140.csv")                        #final_bacteria_ibs_140
  
  fbh_stats <- openxlsx::read.xlsx("./data/final_health_statistic.xlsx")        #final_health_statistic
  fb140_stats <- openxlsx::read.xlsx("./data/final_ibs_140_statistic.xlsx")     #final_ibs_140_statistic

```


## Custom funcs

```{r funcs}

pivot_long_foo <- function(dt){
    dt %>% 
      pivot_longer(
        cols = X01D2Z36_F:possible.genus.03_G,
        names_to = "Bacteria_tax",
        values_to = "TaxonAbun"
      )
}
  
run_wilcox <- function(dat ,cat) {
    dat %>% 
      filter(Category == cat) %>% 
      group_by(TaxonName) %>% 
      summarise(p.value = wilcox.test(TaxonAbun ~ Value)$p.value) %>%
      mutate(Category = cat)
} 

run_kruskal <- function(dat ,cat) {
    dat %>% 
      filter(Category == cat) %>% 
      group_by(TaxonName) %>% 
      summarise(p.value = kruskal.test(TaxonAbun ~ Value)$p.value) %>%
      mutate(Category = cat)
} 

NA_count_foo <- function(dt) {
  dt %>%
    mutate(n = n()) %>%
    summarise(across(everything(), ~ sum(is.na(.x))),
              n = n()) %>%
    select(n, where(function(x) x != 0)) %>%
    mutate(across(everything()), round(. / n, 4)) %>%
    select(-n) %>%
    pivot_longer(
      cols = everything(),
      names_to = "Feature",
      values_to = "Value"
    ) %>%
    arrange(desc(Value))
}

```

## Data processing, step 1

### 'bac_gr_funcs' dataset 

```{r processing1, echo=FALSE}

### Pathogenic and opportunistic bacteria
  bac_gr_funcs1 <- bac_gr_funcs1 %>% 
    mutate(Inflammatory = coalesce(Inflammatory, 0),
           Oral = coalesce(Oral, 0),
           Gases = coalesce(Gases, 0),
           TaxonName = gsub(" ", "_", TaxonName),
           Bacteria_tax = paste(TaxonName, Rank, sep = "_"),
           PATHFL = "Y") %>% 
    select(-Rank) 

### Serotonin producers
  bac_gr_funcs2 <- bac_gr_funcs2 %>% mutate(Destroy = coalesce(Destroy, 0),
                                            SERFL = "Y") %>% select(-Rank)
### Probiotics
  # bac_gr_funcs3 <-
  
### Bacteria with special characteristics
  bac_gr_funcs4 <- bac_gr_funcs4 %>% mutate(SPFL = "Y") %>% select(-Rank)
  
### Vitaminous
  # bac_gr_funcs5 <- 

### SCFA producers
  bac_gr_funcs6 <- bac_gr_funcs6 %>% mutate(FAFL = "Y") %>% select(-Rank)
  bac_gr_funcs6 <- rows_append(bac_gr_funcs6, 
                     tibble(TaxonName = "Ruminococcaceae", 
                     `Ацетат` = 0,
                     `Пропионат` = 0, 
                     Масляная.кислота = 1, 
                     FAFL = "Y"))
  bac_gr_funcs6 <- bac_gr_funcs6 %>% mutate(TaxonName = gsub("\\(.*\\)", "", TaxonName))
  bac_gr_funcs7 <- bac_gr_funcs7 %>% mutate(CMFL = "Y") %>% select(-Rank)
  
### Unhealthy habits  
  bac_gr_funcs7 <- bac_gr_funcs7 %>% pivot_wider(names_from = Habbit, values_from = Habit_state)

```

### 'fb140/fb140_stats/fbh/fbh_stats' dataset 

```{r processing2}

## make dts long

  fb140_long <- pivot_long_foo(fb140)
  fbh_long <- pivot_long_foo(fbh)

## merge dts(fb40_long/fbh_long with corresponding stats dts)

  fb140_mrg <- fb140_long %>% 
    inner_join(fb140_stats, by="patient_ID") 
  
  fbh_mrg <- fbh_long %>% 
    inner_join(fbh_stats, by="patient_ID") 
  
## set fb40/fbh dts together 

  fb_long <- bind_rows(fbh_mrg, fb140_mrg) %>% 
    mutate(Rank = str_split_i(Bacteria_tax, "_", 2)) %>%
    filter(Rank == "F")
  
```

### Final merge of all dts 

```{r merge}

  fb_long. <- fb_long %>% 
    left_join(bac_gr_funcs1, by="Bacteria_tax") %>% 
    mutate(TaxonName = str_split_i(Bacteria_tax, "_", 1)) 

  fb_all0 <- plyr::join_all(list(fb_long., bac_gr_funcs2, bac_gr_funcs4, bac_gr_funcs6, bac_gr_funcs7),
                                by="TaxonName",
                                type="left")

```

## Data processing, step 2

```{r processing3}

 fb_all0 %>%
   map(function(x) sum(is.na(x)) / length(x)) %>%
   enframe() %>%
   unnest(cols = value) %>%
   arrange(desc(value)) %>% 
   filter(name != "Age" & value > 0.26) %>% 
   select(name) %>% 
   pull() -> NA_cols

  NA_count_foo(fb_all0)

  fb_all0.1 <- fb_all0 %>% 
    select(- NA_cols) %>% 
    select(-c("Instrument", "Isolation_source", "Assay_type", "Target_gene", "Rank")) %>% #remove cols with constant value
    mutate(Age_min_max = if_else(!is.na(Age_min), paste(Age_min, Age_max, sep="-"), NA_character_))
 
# sex among researches  
  
  fb_all0.1 %>% 
    filter(Sex %in% c("male", "female")) %>%
    group_by(research_ID, Sex) %>% 
    dplyr::summarise(sumPct = sum(TaxonAbun)) %>%
    ggplot(aes(x=research_ID, y=sumPct, fill=Sex))+
    geom_bar(position="dodge", stat="identity")

# country among researches  

  fb_all0.1 %>% 
    group_by(research_ID, Country) %>% 
    dplyr::summarise(sumPct = sum(TaxonAbun)) %>%
    ggplot(aes(x=research_ID, y=sumPct, fill=Country))+
    geom_bar(position="dodge", stat="identity")

# derive dummy cols    
  
  fb_all1 <- fb_all0.1 %>% 
    mutate(SexUniFL = 
             case_when(research_ID %in% c("research_1", "research_4") ~ 1,
                       research_ID %in% c("research_2", "research_3") ~ 0,
                       .default = NA),
           CountryUniFL = 
             case_when(research_ID == "research_3" ~ 0,
                       .default = 1),
           YouthFL = 
             case_when(!is.na(Age) & between(Age, 15, 24) ~ 1,  
                        is.na(Age) & Age_min_max %in% c("16-42", "18-40", "21-43", "23-28") ~ 1,
                      .default = 0),
           AdultFL = 
             case_when(!is.na(Age) & between(Age, 25, 64) ~ 1,  
                       is.na(Age) & Age_min_max %in% c("16-42", "18-40", "21-43", "23-28", "28-50") ~ 1,
                       .default = 0),
           SeniorFL = 
             case_when(!is.na(Age) & Age >= 65 ~ 1,  
                       .default = 0),
           OneSeqRegionFL = ifelse(stringr::str_detect(Seq_region, "-"), 0, 1))

```

```{r graphs, fig.height=6, fig.width=7, warning=F}

# unique taxons 
  
fb_all1 %>% 
  group_by(research_ID) %>% 
  distinct(TaxonName) -> uniqTax 

ggplot(data=fb_all1, aes(x=research_ID, y=TaxonAbun, color=research_ID, shape=Health_state))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter()+
  ggtitle("Distribution of taxon percentages across Research_ID") +
  labs(y = "Taxon Abundance (%)", x="Research_ID")+
  theme(plot.title = element_text(hjust = 0.5))+
  guides(colour = "none")

# ggsave(filename = file.path("pics","fig1_family.png"))

ggplot(data=uniqTax, aes(x=research_ID, fill=research_ID)) +
  geom_bar()+
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  ggtitle("Count of unique taxa by Research_ID") +
  labs(y = "Count of unique taxa", x="Research_ID")+
  theme(legend.position = "None", plot.title = element_text(hjust = 0.5))

# ggsave(filename = file.path("pics","fig2_family.png"))

```

## Analysis

### Gas-producing bacteria

```{r gas}

  gazbac <- bac_gr_funcs1 %>% filter(Gases == 1) %>% select(TaxonName) %>%  pull()

  fb_all1.gas <- fb_all1 %>% 
    filter(TaxonName == gazbac) %>% 
    mutate(DISFL = if_else(Health_state == "Health", 0, 1)) %>% 
    select(patient_ID, research_ID, Health_state, DISFL, TaxonAbun)
  
  ggplot(data=fb_all1.gas, aes(x=Health_state, y=TaxonAbun, fill=Health_state)) +
    geom_boxplot(outlier.shape = NA)+
    geom_jitter() +
    labs(x="Health state", y="Taxon Abundance (%)") +  
    ggtitle("Veillonellaceae distribution among Health state") +
    theme(legend.position = "None", plot.title = element_text(hjust = 0.5))
  ggsave(filename = file.path("pics","fig2.5_family.png"))
           
    rstatix::wilcox_test(TaxonAbun ~ DISFL, data = fb_all1.gas) %>%   flextable::flextable() %>% flextable::theme_box()
  
  logit1 <- glm(DISFL ~ TaxonAbun, family = "binomial", data = fb_all1.gas)
  summary(logit1)
  
  broom::tidy(logit1) %>% 
    mutate(exp_estimate = exp(estimate),
           p.value = round(p.value, 4)) %>% 
    relocate(exp_estimate, .after = estimate) %>%
    mutate(across(where(is.numeric), function(x) x %>% round(., 4))) %>% 
    flextable::flextable() %>% 
    flextable::theme_box()
    

  ggplot(data=fb_all1.gas, aes(y = DISFL, x = TaxonAbun))+
    geom_point(alpha=.5)+
    stat_smooth(method = "glm", method.args = list(family="binomial"), se = FALSE, color="navy") + 
    labs(x="Taxon Abundance (%)", y="DiseaseFL") +  
    ggtitle("Veillonellaceae") +
    theme(plot.title = element_text(hjust = 0.5))
  
# ggsave(filename = file.path("pics","fig3_family.png"))

```
#### Brunner-Munzel

```{r brunnermunzel}

broom::tidy(brunnermunzel.test(TaxonAbun ~ DISFL, data = fb_all1.gas)) %>% 
  flextable::flextable() %>% flextable::theme_box()

```


### Wordcloud
#### Sex

```{r wordcloud-sex}

  fb_all1.sex <- fb_all1 %>% filter(Sex %in% c("female", "male"))
  fb_all1.sex_long <- fb_all1.sex %>% gather(key = "Category", value = "Value", -c(TaxonName, TaxonAbun))

  map_df("Sex", ~run_wilcox(fb_all1.sex_long, .x)) %>%
    mutate(p.adj = p.adjust(p.value, method = "BH")) %>%
    filter(!is.na(p.adj) & p.adj < 0.05) %>%
    arrange(Category, p.adj) -> rslts.sex 
  
  showbac.sex <- fb_all1.sex %>%
    filter(TaxonName %in% rslts.sex$TaxonName) %>% 
    select(TaxonAbun, TaxonName, Sex) %>% 
    group_by(Sex, TaxonName) %>% 
    summarise(meanTaxonAbun = mean(TaxonAbun)*100) %>% 
    filter(meanTaxonAbun > 0) %>% 
    arrange(meanTaxonAbun) %>% 
    filter(meanTaxonAbun > quantile(meanTaxonAbun, 0.05) & meanTaxonAbun < quantile(meanTaxonAbun, 0.95)) %>% 
    ungroup() %>% 
    rename(word = TaxonName, freq = meanTaxonAbun) %>% 
    mutate(freq = ceiling(freq)) 
  
  showbac.sex_m <- showbac.sex %>%
    filter(Sex == "male") %>%
    select(-Sex) %>% 
    arrange(desc(freq)) %>% 
    as.data.frame()
  
  showbac.sex_f <- showbac.sex %>%
    filter(Sex == "female") %>%
    select(-Sex) %>% 
    arrange(desc(freq)) %>% 
    as.data.frame()
  
  # jpeg(file="./pics/fig4_family.jpeg", width=800, height=800)
  wordcloud(words = showbac.sex_m$word, freq = showbac.sex_m$freq, min.freq = 1,
          max.words=35, random.order=FALSE, rot.per=0.35,
          colors=brewer.pal(8, "Dark2"), scale=c(6, .9))
  # dev.off()
  
  # jpeg(file="./pics/fig5_family.jpeg", width=800, height=800)
  wordcloud(words = showbac.sex_f$word, freq = showbac.sex_f$freq, min.freq = 1,
          max.words=35, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"), scale=c(8, .9))
  # dev.off()

```

#### HealthState

```{r wordcloud-health-state}
  
  fb_all1.hs <- fb_all1
  fb_all1.hs_long <- fb_all1.hs %>% gather(key = "Category", value = "Value", -c(TaxonName, TaxonAbun))

  map_df("Health_state", ~run_wilcox(fb_all1.hs_long, .x)) %>%
    mutate(p.adj = p.adjust(p.value, method = "BH")) %>%
    filter(!is.na(p.adj) & p.adj < 0.05) %>%
    arrange(Category, p.adj) -> rslts.hs

  showbac.hs <- fb_all1.hs %>%
    filter(TaxonName %in% rslts.hs$TaxonName) %>%
    select(TaxonAbun, TaxonName, Health_state) %>%
    group_by(Health_state, TaxonName) %>%
    summarise(meanTaxonAbun = mean(TaxonAbun)*100) %>%
    filter(meanTaxonAbun > 0) %>%
    arrange(meanTaxonAbun) %>%
    filter(meanTaxonAbun > quantile(meanTaxonAbun, 0.05) & meanTaxonAbun < quantile(meanTaxonAbun, 0.95)) %>%
    ungroup() %>% 
    rename(word = TaxonName, freq = meanTaxonAbun) %>% 
    mutate(freq = ceiling(freq)) 
  
  showbac.hs_h <- showbac.hs %>%
    filter(Health_state == "Health") %>%
    select(-Health_state) %>% 
    arrange(desc(freq)) %>% 
    as.data.frame()
  
  showbac.hs_d <- showbac.hs %>%
    filter(Health_state == "Disease") %>%
    select(-Health_state) %>% 
    arrange(desc(freq)) %>% 
    as.data.frame()
  
  # jpeg(file="./pics/fig6_family.jpeg", width=800, height=800)
  wordcloud(words = showbac.hs_h$word, freq = showbac.hs_h$freq, min.freq = 1,
          max.words=35, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"), scale=c(6, .9))
  # dev.off()
  
  # jpeg(file="./pics/fig7_family.jpeg", width=800, height=800)
  wordcloud(words = showbac.hs_d$word, freq = showbac.hs_d$freq, min.freq = 1,
          max.words=35, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"), scale=c(6, .9))
  # dev.off()
  
```

<!-- ### Mixed effects -->
<!-- #### research_ID -->

```{r}
# 
# fb_all1.mix <- fb_all1 %>% 
#     mutate(across(c(Sex, patient_ID, research_ID, 
#                     TaxonName, Health_state), function(x) x %>% factor()))

# fb_all1.mix <- fb_all1 %>% 
#      mutate(research_ID = factor(research_ID), 
#             TaxonAbun_d100 = TaxonAbun/100)
# 
#  mod_rslts.resID <- list()
# 
#   for(taxon in unique(fb_all1.mix$TaxonName)) {
#     
#     taxon_data <- fb_all1.mix %>% filter(TaxonName == taxon) 
#     
#     if (sum(taxon_data$TaxonAbun_d100) > 0) {
#       model <- glmmTMB(TaxonAbun_d100 ~ research_ID + (1 | research_ID), data = taxon_data,
#                        family = beta_family(link="logit"),  zi=~1 )
#       mod_rslts.resID[[taxon]] <- summary(model)
#     }
#   }
 
```

## ML (random forest)

### preprocessing

```{r processing4}

cpk2_raw0 <- fb_all1 

cpk2_raw0 %>% glimpse()

cpk2_raw1 <- cpk2_raw0 %>% 
  select(-c("Age", "Age_min", "Age_max", "Age_min_max", "Seq_date", 
            "Sex", "SexUniFL", "AdultFL", "SeniorFL", "Bacteria_tax", "Country",
            "CountryUniFL", "Seq_region"))

cpk2.1 <- cpk2_raw1 %>% 
  select(patient_ID, research_ID, TaxonName, TaxonAbun) %>% 
  pivot_wider(names_from = TaxonName, values_from = TaxonAbun)

cpk2.2 <- cpk2_raw1 %>% 
  select(-TaxonName, -TaxonAbun) %>% 
  distinct()

cpk2.2 %>% 
  inner_join(cpk2.1, by = join_by(patient_ID, research_ID)) %>% 
  mutate(across(ends_with("FL"), function(x) x %>% factor(levels = c("0", "1"))),
         HasDiseaseFL = factor(ifelse(Health_state == "Disease", 1, 0), levels=c("0", "1"))) %>% 
  relocate(HasDiseaseFL) -> cpk2.2.5

cpk2.2.5 %>% 
  select(-patient_ID, -Health_state) -> cpk2.3

```

### Clustering (Partitioning Around Medoids, variable = research_ID) 

```{r cluster}

cpk2.2.5 %>% select(-patient_ID, -Health_state, -HasDiseaseFL, -YouthFL, -OneSeqRegionFL) -> cpk_clust     #research_ID, HasDiseaseFL

# cpk2.2.5 %>% select(-patient_ID, -Health_state) -> cpk_clust     
#options = research_ID, HasDiseaseFL

cpk_clust.1 <- cpk_clust[,-1] %>% select_if(colSums(.) != 0)

fit.pam <- pam(cpk_clust.1, k=6, stand = F, metric = "manhattan")

ct.pam <- table(cpk_clust$research_ID, fit.pam$clustering)
# ct.pam

randIndex(ct.pam) # adjusted Rand index

fviz_cluster(fit.pam, data=cpk_clust.1)
# ggsave(filename = file.path("pics","fig8_family.png"))
```

### Boruta

```{r boruta}

# NA_count_foo(cpk2.3)

cpk2.3 <- cpk2.3 %>% select(-research_ID)

Boruta(HasDiseaseFL ~ ., cpk2.3, ntree = 500, maxRuns = 500) %>%
  TentativeRoughFix() -> boruta_trained

boruta.sig.vars <- boruta_trained %>%
  attStats() %>%
  filter(decision == "Confirmed" & meanImp > 5) %>% 
  arrange(desc(meanImp)) %>% 
  rownames_to_column(var = "ImpVar") %>% 
  select(ImpVar) %>% 
  pull()

boruta_trained %>%
  attStats() %>%
  filter(decision == "Confirmed") %>%  
  rownames_to_column("Variable") %>%
  mutate(Variable = Variable %>% fct_reorder(`meanImp`)) %>%
  
  ggplot(aes(y = Variable, x = meanImp, colour = decision)) +
  geom_point() +
  geom_errorbar(aes(xmin = minImp, xmax = maxImp, width = 0.1)) +
  xlab("Entropy mean reduction") +
  labs(color = "Variable significance") +
  theme(legend.position = "bottom")
  # ggsave(filename = file.path("pics","fig9_family.png"), dpi = 300)

```

### make train/test dts

```{r train-test, warning=F}

cpk2.4 <- cpk2.3[sample(nrow(cpk2.3)), ]

cpk2 <- cpk2.4 %>% select(HasDiseaseFL, boruta.sig.vars)

split_train_test <- initial_split(cpk2, strata = HasDiseaseFL, prop = 0.80)

cpk2_train <- split_train_test %>% training()
cpk2_test <- split_train_test %>% testing()

```

### Corrplot

```{r, fig.width=10, fig.height=10}

cpk2 %>%
  select(where(is.numeric) & !HasDiseaseFL) %>%
  corr.test() %>%
  .$r %>%
  ggcorrplot(lab = TRUE)

```

### Recipe

#### Recipe, step1

```{r recipe1}

cat_metric <- yardstick::metric_set(
  
    yardstick::bal_accuracy,
    yardstick::precision,
    yardstick::recall,
    yardstick::f_meas,
    yardstick::sensitivity,
    yardstick::specificity,
    yardstick::j_index
  
  )

```

#### Recipe, step2

```{r recipe2}

# cpk2_recipe <- recipe(HasDiseaseFL ~ ., cpk2_train) %>%
#   step_zv(all_predictors()) %>%                   #zero variance vars 
#   step_nzv(all_predictors()) %>%                  #non-balanced vars
#   step_lincomb(all_numeric_predictors()) %>%      #linear combinations (with high correlation coef)
#   # step_normalize(all_numeric_predictors()) %>%
#   step_lencode_glm(all_nominal_predictors(), outcome = vars(HasDiseaseFL)) %>% #target encoding
#   step_adasyn(HasDiseaseFL)  %>%               #compensation of Disease/Health disbalance
#   prep()
# 
# cpk2_recipe %>% juice() %>% count(HasDiseaseFL)
# cpk2_train %>% count(HasDiseaseFL)
 
cpk2_recipe <- recipe(HasDiseaseFL ~ ., cpk2_train) %>%
  step_zv(all_predictors()) %>%
  step_nzv(all_predictors()) %>%
  step_lincomb(all_numeric_predictors()) %>%
  step_lencode_glm(all_nominal_predictors(), outcome = vars(HasDiseaseFL)) %>%
  step_adasyn(HasDiseaseFL) #%>%
  #prep()
```

### Model definition

```{r define-model}
rf_model <- rand_forest(mode = "classification", mtry = tune(), trees = tune(), min_n = tune()) %>%
  set_engine("ranger")
```

### Sample definition

```{r define-sample}
cv_samples <- vfold_cv(cpk2_train, strata = HasDiseaseFL, v = 5)
```

### Parameter grid definition

```{r define-param-grid}
parameters_grid <- grid_max_entropy(mtry(range = c(5, 12)), trees(), min_n(), size = 20)
```

### Pipeline definition

```{r define-pipe}

reg_workflow <- workflow() %>%
  add_recipe(cpk2_recipe) %>%
  add_model(rf_model)

```

### Model training

```{r train-model}

grid_search <- reg_workflow %>%
  
  tune_grid(
    
    object = reg_workflow,
    resamples = cv_samples,
    grid = parameters_grid,
    control = control_grid(save_pred = TRUE),

    metrics = metric_set(sensitivity, specificity, j_index)

  )

```

### Visualization

```{r visualize-model}

grid_search %>%
  collect_metrics() %>%
  ggplot(aes(trees, mean, color = .metric)) +
  geom_errorbar(aes(ymin = mean - std_err, ymax = mean + std_err), alpha = 0.8) +
  geom_line(size = 1)

```

### Best model choice

```{r best-model}

best_by_j_index <- grid_search %>% select_best("j_index")

# best_by_j_index %>%   flextable::flextable() %>% flextable::theme_box()

final_reg_model <- finalize_workflow(
  
  reg_workflow,
  best_by_j_index
  
)

```

## Compare metrics on test and train data

```{r metrics-compare}
# test data

final_reg_model %>%
  keras::fit(cpk2_train) %>%
  predict(cpk2_test) %>%
  pull() -> final_test_prediction

metrics_on_test <- cat_metric(truth = truth_values, estimate = estimate_values, tibble(truth_values = cpk2_test$HasDiseaseFL, estimate_values = final_test_prediction)) %>% rename(test_estimate = `.estimate`) %>% select(!`.estimator`)

# train data

final_reg_model %>%
  keras::fit(cpk2_train) %>%
  predict(cpk2_train) %>%
  pull() -> final_train_prediction

metrics_on_train <- cat_metric(truth = truth_values, estimate = estimate_values, tibble(truth_values = cpk2_train$HasDiseaseFL, estimate_values = final_train_prediction)) %>% rename(train_estimate = `.estimate`) %>% select(!`.estimator`)

# binding

metrics_on_test %>%
  left_join(metrics_on_train, by = ".metric") %>%
  mutate(differencies = train_estimate - test_estimate) %>% 
  # write.csv(.,file = "./output/metrics.csv")
  flextable::flextable() %>% flextable::theme_box()
```

### Final estimation

```{r final-estimation}

last_fit(
  
  final_reg_model,
  split_train_test
  
) -> final_log_model_all_data

final_log_model_all_data %>%
  extract_workflow() %>%
  predict(cpk2, type = "class") %>%
  pull() -> class_prediction

final_log_model_all_data %>%
  extract_workflow() %>%
  predict(cpk2, type = "prob") %>%
  pull() -> prob_prediction

log_model_results <- tibble(truth = cpk2$HasDiseaseFL,
                            estimate = class_prediction,
                            prob_yes = 1 - prob_prediction)

cat_metric(truth = truth, estimate = estimate, log_model_results) 

log_model_results %>%
  roc_curve(truth = truth, prob_yes) %>%
  autoplot()
# ggsave(filename = file.path("pics","fig10_family_1.png"))
```
```{r}

```

## DBSCAN

```{r dbscan}

cpk2.3 ->  cluster_data_for_vizualisation

cpk2.3 %>% select(-HasDiseaseFL) -> cluster_data_for_model

recipe(~ ., data = cluster_data_for_model) %>%
  step_zv(all_predictors()) %>%
  step_scale(all_numeric_predictors()) %>%
  #step_dummy(Sex) %>%
  prep(cluster_data_for_model) %>%
  juice() -> data_recoded

data_recoded %>%
  daisy(metric = "gower") %>%
  umap(n_components = 10) -> umap_data

data_recoded %>%
  daisy(metric = "gower") %>%
  umap(n_components = 3) -> umap_data_for_visualization

umap_data %>%
  daisy(metric = "gower") -> dist_data

dist_data %>%
 dbscan::dbscan(eps = mean(dist_data)) -> cluster_data_for_model

# dist_data %>%
#   dbscan::hdbscan(minPts = 5) -> cluster_data_for_model

umap_data_for_visualization %>%
  as_tibble() %>%
  mutate(Cluster = as.factor(str_c("Cluster №", cluster_data_for_model$cluster)),
         HasDiseaseFL = cluster_data_for_vizualisation$HasDiseaseFL) %>%
  rename( Component1 = V1,
          Component2 = V2,
          Component3 = V3) -> data_3d

umap_fig <- plot_ly(data_3d,
        x =~ Component1,
        y =~ Component2,
        z =~ Component3,
        color =~ Cluster, hoverinfo = "text", text = ~paste(HasDiseaseFL)) %>%
  
  add_markers(size = 1.5)

plot_ly(data_3d,
        x =~ Component1,
        y =~ Component2,
        z =~ Component3,
        color =~ Cluster, hoverinfo = "text", text = ~paste(HasDiseaseFL)) %>%
  
  add_markers(size = 1.5)

htmlwidgets::saveWidget(
                widget = umap_fig, #the plotly object
                file = "./pics/umap_figure.html", #the path & file name
                selfcontained = TRUE #creates a single html file
                )

```
